Extracting Angular Observables without a Likelihood\\and Applications to Rare Decays

### 2015-03-11 draft.tex
Update: rearrange plots, rewrite text on fit vs. mom estimators

Our findings can be summarized as follows:
\begin{itemize}
    \item In all studied cases we observed not a single bias in the distribution of the $\mathrm{pull}$
        of any observables $S_i$,
        \begin{equation}
            \mathrm{pull_i} \equiv \frac{\est{S}_i - S_i}{\est{\sigma}_{i}}\,.
        \end{equation}
        Here $S_i$ refers to the true (input) value for the angular observables, $\est{S}_i$ refers to the
        mean of the pseudo measurement via the MC estimate, and $\est{\sigma}_{i} \equiv \est{\Sigma_{ii}}^{1/2}$
        refers to an estimator of the standard deviation of the pseudo measurement. All pull distributions
        obtained in our studies can be successfully
        fitted to a Gaussian distribution. Out of the large number  of studied distributions, we only show the pull distributions for the observables
        $S_5 \simeq 27\%$ and $S_7 \simeq 2\%$ obtained from SM-like $B\to K^*\ell^+\ell^-$
        decays as representative examples. We generated $2\cdot 10^5$ toy studies
        with $200$ events per study in \reffig{pulls}.
    \item We study the MC estimate for the absolute uncertainty $\est{\sigma_i(N)}$ with respect to the angular observable $S_i$
        as a function of the number of simulated events $N$. As expected for a multivariate Gaussian
        distribution, we find that the absolute uncertainty is well fitted by
        \begin{equation}
            \label{eq:unc-on-mean}
            \est{\sigma_i(N)} = \frac{\sigma_i}{\sqrt{N}}
        \end{equation}
        with $\sigma_i(1) = \order{1}$, regardless of the absolute size of $S_k$. The latter can best be shown
        for the example of uncertainties of two observables. Taking again $S_5$ ($\simeq 27\%$) and $S_7$ ($\simeq 2\%$)
        for SM-like $B\to K^*\ell^+\ell^-$ decays, we show the absolute uncertainty in \reffig{errors}.
        We find that the the method of moments yields uncertainties on $\vec{S}$ that are roughly
        $10\%$ -- $15\%$ larger than those obtained from maximum-likelihood fits and for the same
        number of events \danny{[cite!]}. However, we wish to note that said fits only produce a limited
        subset of the angular observables, and the statistical error of the fit is expected to increase
        with the number of fit parameters until their errors saturate the statistical errors of the method-of-moments
        estimators \cite[sec. 8]{Cowan:1998ji}.
%    \item For all toy studies, we produce plots of their skewness and their kurtosis as functions of
%        $N$. We find in all cases that the mean skewness and kurtosis converge towards zero.
%        \todo{Plot Skewness and Kurtosis for $S_5$ and $S_7$ as functions of the number of simulated events $N$.}
\end{itemize}
\begin{figure}[b]
        \centering
            \includegraphics[width=0.45\textwidth]{figs/pull-Q2_5_6_S5_200.png}
            \includegraphics[width=0.45\textwidth]{figs/pull-Q2_5_6_S7_200.png}
        \caption{Pull distribution for the angular observables $S_5$ (left) and $S_7$ (right), extracted from $2\cdot 10^5$ studies of
        $200$ simulated events each for the decay $B\to K^*\ell^+\ell^-$. The red curve represents a fit to a Gaussian distribution.}
        \label{fig:pulls}
\end{figure}

@@

\begin{figure}[t]
        \centering
            \includegraphics[width=0.45\textwidth]{figs/Q2_1_2_S4.png}
            \includegraphics[width=0.45\textwidth]{figs/Q2_1_2_S8.png}
        \caption{Uncertainty of the angular observables $S_4$ (left) and $S_8$ (right), extracted from $2\cdot 10^5$ studies of
        simulated events for the decay $B\to K^*\ell^+\ell^-$. We show the uncertainty as a function of the number of simulated events $N$.
        The red curve represents a fit to the function given in \refeq{unc-on-mean}.}
        \label{fig:errors}
\end{figure}

We also studied the correlation of the estimators from Method of Moments to a maximum likelihood fit. We performed a toy study by simulating $10^3$ toys of 200 events. We then use for each toy method described in~\refsec{systematics:partial-waves} and a standard maximum likelihood method. In \reffig{correlation} the two estimators distributions are showed. In conclusion the two estimators are highly correlated but different between then can be of the order of statistic uncertainty of the maximum likelihood estimator. We found that in some observables this can reach almost half of the statistical error.

\begin{figure}[t]
        \centering
            \includegraphics[width=0.45\textwidth]{figs/S3_scat.pdf}
            \includegraphics[width=0.45\textwidth]{figs/S5_scat.pdf}
        \caption{Correlation between Method of Moments and maximum likelihood estimators for observables $S_3$ and $S_5$. }
        \label{fig:correlation}
\end{figure}

>>>>

\begin{figure}[b]
        \centering
            \includegraphics[width=0.45\textwidth]{figs/pull-Q2_5_6_S5_200.png}
            \includegraphics[width=0.45\textwidth]{figs/pull-Q2_5_6_S7_200.png}
        \caption{Pull distribution for the angular observables $S_5$ (upper) and $S_7$ (lower), extracted from $2\cdot 10^5$ studies of
        $200$ simulated events each for the decay $B\to K^*\ell^+\ell^-$. The red curve represents a fit to a Gaussian distribution.}
        \label{fig:pulls}
\end{figure}

Our findings can be summarized as follows:
\begin{itemize}
    \item In all studied cases we observed not a single bias in the distribution of the $\mathrm{pull}$
        of any observables $S_i$,
        \begin{equation}
            \mathrm{pull_i} \equiv \frac{\est{S}_i - S_i}{\est{\sigma}_{i}}\,.
        \end{equation}
        Here $S_i$ refers to the true (input) value for the angular observables, $\est{S}_i$ refers to the
        mean of the pseudo measurement via the MC estimate, and $\est{\sigma}_{i} \equiv \est{\Sigma_{ii}}^{1/2}$
        refers to an estimator of the standard deviation of the pseudo measurement. All pull distributions
        obtained in our studies can be successfully
        fitted to a Gaussian distribution. Out of the large number  of studied distributions, we only show the pull distributions for the observables
        $S_5 \simeq 27\%$ and $S_7 \simeq 2\%$ obtained from SM-like $B\to K^*\ell^+\ell^-$
        decays as representative examples. We generated $2\cdot 10^5$ toy studies
        with $200$ events per study in \reffig{pulls}.

    \item We study the MC estimate for the absolute uncertainty $\est{\sigma_i(N)}$ with respect to the angular observable $S_i$
        as a function of the number of simulated events $N$. As expected for a multivariate Gaussian
        distribution, we find that the absolute uncertainty is well fitted by
        \begin{equation}
            \label{eq:unc-on-mean}
            \est{\sigma_i(N)} = \frac{\sigma_i}{\sqrt{N}}
        \end{equation}
        with $\sigma_i(1) = \order{1}$, regardless of the absolute size of $S_k$. The latter can best be shown
        for the example of uncertainties of two observables. Taking again $S_5$ ($\simeq 27\%$) and $S_7$ ($\simeq 2\%$)
        for SM-like $B\to K^*\ell^+\ell^-$ decays, we show the absolute uncertainty in \reffig{errors}.
        We find that the the method of moments yields uncertainties on $\vec{S}$ that are roughly
        $10\%$ -- $15\%$ larger than those obtained from maximum-likelihood fits and for the same
        number of events \danny{[cite!]}. However, we wish to note that said fits only produce a limited
        subset of the angular observables, and the statistical error of the fit is expected to increase
        with the number of fit parameters until their errors saturate the statistical errors of the method-of-moments
        estimators \cite[sec. 8]{Cowan:1998ji}.

    \item We also compare the results as obtained by the method of moments with
        results obtained by a conventional likelihood fit. In particular, we
        study the correlation between the method-of-moment estimators and the
        maximum-likelihood-fit estimators. We run $10^3$ toy analyses, with
        $200$ simulated events per analysis. We show the joint distribution of
        the two estimators in \reffig{correlation}. The two estimators are
        highly correlated. The distribution of the difference of the estimators
        exhibits now bias, which is due to the large number of simulated
        events.  Still, we find that statistical uncertainty on the difference
        of the two estimators is sizeable and can easily become half as large
        as the statistical uncertainty of either estimator.
\end{itemize}
\begin{figure}[t]
        \centering
            \includegraphics[width=0.45\textwidth]{figs/Q2_1_2_S4.png}
            \includegraphics[width=0.45\textwidth]{figs/Q2_1_2_S8.png}
        \caption{Uncertainty of the angular observables $S_4$ (uppwe) and $S_8$
            (lower), extracted from $2\cdot 10^5$ studies of simulated events
            for the decay $B\to K^*\ell^+\ell^-$. We show the uncertainty as a
            function of the number of simulated events $N$.  The red curve
        represents a fit to the function given in \refeq{unc-on-mean}.}
        \label{fig:errors}
\end{figure}

\begin{figure}[t]
        \centering
            \includegraphics[width=0.45\textwidth]{figs/S3_scat.pdf}
            \includegraphics[width=0.45\textwidth]{figs/S5_scat.pdf}
            \caption{The joint distribution of the two estimators
                $\est{S}_i^\text{MoM}$ and $\est{S}_i^\text{Fit}$ that arise
                from the method of moments and a maximum-likelihood fit,
                respectively. We show the estimators for both of our benchmark
                observables $S_3$ (upper) and $S_5$ (lower).
                \danny{Marcin: please replace by $S_5$ (upper) and $S_7$ (lower).}
            }
        \label{fig:correlation}
\end{figure}

@@
